---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and pre-processing the data

Load dplry, ggplot2, knitr, and mice libraries:
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(mice)
```

Unzip file to access CSV and read file: 
```{r}
if(!file.exists("activity.csv")){unzip("activity.zip")}
activity<-read.csv("activity.csv", stringsAsFactors = FALSE, na.strings="NA")
```

## What is mean total number of steps taken per day?  

Group data by date and find mean and median for each day: 
```{r}
  steps_per_day<-activity %>%
    group_by(date) %>%
    summarise(mean_steps = mean(steps, na.rm = TRUE),
            median_steps = median(steps, na.rm = TRUE))
  kable(steps_per_day)
```

Calculate total steps per day and create histogram: 
```{r}
  by_date<-group_by(activity, date)
  total_steps<-summarise(by_date, daily_steps=sum(steps, na.rm = TRUE))
  hist(total_steps$daily_steps, breaks=30, xlab="Total Steps", main="Total Steps per Day")
```

## What is the average daily activity pattern?
Group data by interval and find mean for each interval: 
```{r}
daily_pattern<-activity %>%
  group_by(interval) %>%
  summarise(interval_steps = mean(steps, na.rm = TRUE))
```

Plot average steps taken per interval: 
```{r}
plot(daily_pattern$interval, daily_pattern$interval_steps, type = "l", xlab="Interval", ylab="Average Steps", main="Average Steps per Interval")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
daily_pattern[which(daily_pattern$interval_steps==max(daily_pattern$interval_steps)),]
```

## Imputing missing values

Find the total number of missing values in the dataset: 

```{r}
nrow(activity[is.na(activity$steps),])
```

This could also be done using the MICE package: 
```{r}
md.pattern(activity)
```

Use the MICE package to impute missing values using Predictive Mean Matching: 
```{r results="hide"}
imputed_Data <- mice(activity, m=5, maxit = 50, method = 'pmm', seed = 500)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
complete_activity<-complete(imputed_Data,3)
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
comp_total_steps<-complete_activity%>%
    group_by(date)%>%
    summarise(comp_daily_steps=sum(steps, na.rm = TRUE))
  hist(comp_total_steps$comp_daily_steps, breaks=30, xlab="Total Steps", main="Total Steps per Day")

comp_per_day<-complete_activity %>%
    group_by(date) %>%
    summarise(mean_steps = mean(steps, na.rm = TRUE),
            median_steps = median(steps, na.rm = TRUE))
kable(comp_per_day)

```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
complete_activity<-mutate(.data = complete_activity, date=as.Date(date))
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
complete_activity$weekday <- factor((weekdays(complete_activity$date) %in% weekdays), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
```

Make a panel plot containing a time series plot (i.e. 𝚝𝚢𝚙𝚎 = "𝚕") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
weekday_pattern<-complete_activity %>%
  group_by(weekday, interval) %>%
  summarise(interval_steps = mean(steps, na.rm = TRUE))

ggplot(weekday_pattern, aes(x=interval, y=interval_steps, group=weekday, col=weekday)) +
  geom_line() + 
  facet_grid(weekday~.) +
  ggtitle("Activity Pattern of Weekdays and Weekends")
```